---
title: "Chapter 1"
author: "Nicolas May"
date: "January 1, 2019"
output: 
  html_document:
    theme: readable
    highlight: zenburn
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```

# Connect to AWS PostgreSQL Server

The "PostgreSQL" DNS below is defined in `/usr/local/etc/odbc.ini`, which points to `/usr/local/etc/odbcinst.ini` in macOS; in Debian/Ubuntu, they're in `~/.odbc.ini` pointing to `/etc/odbcinst.ini`.

The `dvdrental` database (available as a .tar file at [https://github.com/PacktPublishing/PostgreSQL-11-in-7-days/tree/master/resources](https://github.com/PacktPublishing/PostgreSQL-11-in-7-days/tree/master/resources)) has already been uploaded to the AWS EC2 PostgreSQL server via pgAdmin4's restore functionality. A local version of PostgreSQL must also be installed to use pgAdmin4's restore functionality.

Load useful libraries.
```{r}
library(DBI)
library(odbc)
library(RPostgres)
```

Connecto to the _dvdrental_ database.
```{r}
# "PostgreSQL" DNS set in:
#   /usr/local/etc/odbc.ini
#   /usr/local/etc/odbcinst.ini
# ... OR ...:
#   ~/.odbc.ini
#   /etc/odbcinst.ini
con_dvdrental <- DBI::dbConnect(odbc::odbc(), "PostgreSQL AWS dvdrental")
# RPostgreSQL::postgresqlDescribeConnection(con_dvdrental)
```

# Chapter Sections

## Setup a Practice Database Using Referenced .tar File

See what's in the `customer` table.
```{sql, connection=con_dvdrental}
SELECT * FROM customer;
```

Disconnect from the _dvdrental_ database.
```{r}
dbDisconnect(con_dvdrental)
```


## Create Database, Table, Insert, and Query Data – Part 1

### Creating a database

Here, we'll be using the default `postgres` database (which has PostgreSQL admin tools) to create a new database called `petshop`.

Connect to the _postgres_ database.
```{r}
con_postgres <- DBI::dbConnect(odbc::odbc(), "PostgreSQL AWS postgres")
```

The template for creating a new database is `CREATE DATABASE database_name;`.

```{sql, connection=con_postgres}
-- CREATE DATABASE petshop; -- This database is already created, so it's commented out
```

Disconnect from the `postgres` database.
```{r}
dbDisconnect(con_postgres)
```


## Create Database, Table, Insert, and Query Data – Part 2

Connect to the _petshop_ database.
```{r}
con_petshop <- DBI::dbConnect(odbc::odbc(), "PostgreSQL AWS petshop")
```

### Creating a table

The template for creating a table is `CREATE TABLE table_name(column_name datatype CONSTRAINT);`.

```{sql, connection=con_petshop}
DROP TABLE IF EXISTS dogs; -- if this script is run more than once
CREATE TABLE dogs(
  dog_id serial       PRIMARY KEY, -- serial datatype is auto-incrementing integer
  name   varchar(255) NOT NULL,
  age    integer      NOT NULL,
  weight integer      NOT NULL
);
```

### Inserting data into a table

The template for inserting a row of data into a table is `INSERT INTO table_name(column_1, column_2) VALUE ('value_one', 'value_two');`.

```{sql, connection=con_petshop}
INSERT INTO dogs(
  name, age, weight
) VALUES (
  'Rover', 3, 35
);
```

### Querying data from a table

Run a simple `SELECT * FROM ...` query.

```{sql, connection=con_petshop}
SELECT * FROM dogs;
```

Disconnect from the _petshop_ database.
```{r}
dbDisconnect(con_petshop)
```


## PostgreSQL Relationships Using What’s Called a Primary Key and Foreign Key

The template for connectiong primary and foreign keys is as follows:
```
CREATE TABLE dogs(
  dog_id   serial       PRIMARY KEY,
  owner_id integer,     
    FOREIGN KEY(owner_id) REFERENCES owners(owner_id),
  name     varchar(255) NOT NULL,
  age      integer      NOT NULL,
  weight   integer      NOT NULL
);
```

## Assignment

Connect to _postgres_ database in order to create a _toy_store_ database.
```{r}
con_postgres <- DBI::dbConnect(odbc::odbc(), "PostgreSQL AWS postgres")
```

Create the _toy_store_ database.
```{sql, connection=con_postgres}
-- CREATE DATABASE toy_store; -- This database is already created, so it's commented out
```

Disconnect from the _postgres_ database.
```{r}
dbDisconnect(con_postgres)
```

Now, connect to the _toy_store_ database.
```{r}
con_toy_store <- DBI::dbConnect(odbc::odbc(), "PostgreSQL AWS toy_store")
```

Create the assigned table.
```{sql, connection=con_toy_store}
DROP TABLE IF EXISTS customer;
CREATE TABLE customer(
  customer_id  serial        PRIMARY KEY,
  first_name   varchar(255)  NOT NULL,
  last_name    varchar(255)  NOT NULL,
  email        varchar(255)  NOT NULL
);
```

Insert a row of data into the `customer` table.
```{sql, connection=con_toy_store}
INSERT INTO customer(
  first_name, last_name, email
) VALUES (
  'Bill', 'Sanchez', 'bill_sanchez_666@hotmail.com'
);
```

Query the `customer` table to see if everything worked.
```{sql, connection=con_toy_store}
SELECT * FROM customer;
```

It worked!

Disconnect from the _toy_store_ database.
```{r}
dbDisconnect(con_toy_store)
```


```
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
#  @##==---==##@##==---==##@    EXTRA  :  SPACE    @##==---==##@##==---==##@  #
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
#  @##==---==##@##==---==##@    EXTRA  :  SPACE    @##==---==##@##==---==##@  #
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
```
