---
title: "Chapter 2"
author: "Nicolas May"
date: "1/2/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```

# Connect to AWS PostgreSQL Server

Load useful libraries.
```{r}
library(DBI)
library(odbc)
library(RPostgreSQL)
```

Connecto to the _dvdrental_ database.
```{r}
con_dvdrental <- DBI::dbConnect(odbc::odbc(), "PostgreSQL AWS dvdrental")
```


# Chapter Sections

## Introduction to the SELECT Statement to Retrieve Rows of Data

```{sql, connection=con_dvdrental}
SELECT * FROM customer;
```

```{sql, connection=con_dvdrental}
SELECT first_name, email
  FROM customer;
```

```{sql, connection=con_dvdrental}
SELECT last_name
  FROM customer;
```

```{sql, connection=con_dvdrental}
SELECT country, country_id
  FROM country;
```

What tables are in the _dvdrental_ database?
```{r}
dbListTables(con_dvdrental)
```


## Introduction to the SELECT DISTINCT Statement

```{sql, connection=con_dvdrental}
SELECT * FROM film;
```

```{sql, connection=con_dvdrental}
SELECT DISTINCT rental_rate FROM film;
```

```{r}
dbListFields(con_dvdrental, 'film')
```

```{sql, connection=con_dvdrental}
SELECT DISTINCT rental_duration
  FROM film;
```

```{sql, connection=con_dvdrental}
SELECT DISTINCT rating
  FROM film;
```


## SELECT Statement with WHERE Clause

Here's a template for using the `WHERE` clause to filter results:
```
SELECT column_name
  FROM table_name
  WHERE <condition = true>;
```

Using just a `WHERE` clause:
```{sql, connection=con_dvdrental}
SELECT title, rental_rate
  FROM film
  WHERE rental_rate = 0.99;
```

Using a `WHERE` clause with a logical operator (`AND`, `OR`):
```{sql, connection=con_dvdrental}
SELECT title, rental_rate
  FROM film
  WHERE rental_rate = 0.99
    AND rating = 'PG';
```

```{sql, connection=con_dvdrental}
SELECT first_name, last_name, email
  FROM customer
  WHERE first_name = 'Everett'
    AND last_name  = 'Banda';
```

```{sql, connection=con_dvdrental}
SELECT *
  FROM film
  WHERE title       = 'Boulevard Mob'
    AND rental_rate = 0.99;
```

Out of my own curiosity, what's in the `fulltext` field in the `film` table?
```{r}
full_text_ex <- 
  dbGetQuery(con_dvdrental,
             "SELECT fulltext
                FROM film
                WHERE title = 'Boulevard Mob';")
str(full_text_ex)
full_text_ex <- full_text_ex[[1, 'fulltext']]
class(full_text_ex)
full_text_ex
```

The next section uses the _petshop_ database, so we can close the _dvdrental_ database.
```{r}
dbDisconnect(con_dvdrental)
```


## Insert Data into Tables, Update, and Delete Rows

This section uses the _petshop_ database, so let's create a connection to that.
```{r}
con_petshop <- dbConnect(odbc::odbc(), "PostgreSQL AWS petshop")
dbListTables(con_petshop)
```

```{sql, connection=con_petshop}
SELECT * FROM dogs;
```

### Insert a records using `INSERT INTO`

This next code block simply resets the auto-incrementing `dog_id` field to restart with the number 2.
```{sql, connection=con_petshop}
ALTER SEQUENCE dogs_dog_id_seq RESTART WITH 2;
```

Now for the insert.
```{sql, connection=con_petshop}
INSERT INTO dogs(
  name, age, weight
) VALUES (
  'Spot', 3, 23
);
```

```{sql, connection=con_petshop}
SELECT * FROM dogs;
```

So far, so good.

### Updating a record with `UPDATE`

Now, let's update a record. Let's change Spot's age to 4.
```{sql, connection=con_petshop}
UPDATE dogs
  SET   age    = 4
  WHERE name   = 'Spot';
```

```{sql, connection=con_petshop}
SELECT * FROM dogs;
```

The age update worked.

Now let's change Spot's name to "Sparky".
```{sql, connection=con_petshop}
UPDATE dogs
  SET   name   = 'Sparky'
  WHERE name   = 'Spot';
```

```{sql, connection=con_petshop}
SELECT * FROM dogs;
```

The name update worked.

### Deleting a record with `DELETE`

Now, let's delete the records we just added (`dog_id` 2).

Before we delete, let's first do a `SELECT` ... `WHERE` to make sure we're deleting the right record.
```{sql, connection=con_petshop}
SELECT * 
  FROM dogs
  WHERE name = 'Sparky';
```

That's the one. So, now let's delete it.
```{sql, connection=con_petshop}
DELETE FROM dogs
  WHERE name = 'Sparky';
```

Confirm the deletion.
```{sql, connection=con_petshop}
SELECT * FROM dogs;
```

The record deletion worked.

### Exercise

#### Task 1

Write a query that updates the age column to 5 in the first row. (Really, this isn't a "query", it's a "statement".)

```{sql, connection=con_petshop}
UPDATE dogs
  SET   age    = 5
  WHERE name   = 'Rover';
```

```{sql, connection=con_petshop}
SELECT * FROM dogs;
```

#### Task 2

Write a query that deletes the first row. (Again, "statement", not "query".)

```{sql, connection=con_petshop}
SELECT *
  FROM dogs
  WHERE name = 'Rover';
```

```{sql, connection=con_petshop}
DELETE FROM dogs
  WHERE name = 'Rover';
```

```{sql, connection=con_petshop}
SELECT * FROM dogs;
```

Where done with the _petshop_ database, so we can close the connection.

```{r}
dbDisconnect(con_petshop)
```


## Assignment

### Problem 1

You're an employee at the DVD rental store and someone brings a wallet they found to the front desk. You discover the wallet belongs to Terry Grissom. Using knowledge from section 2, find a way to querry Terry's email from the database to let him know about his wallet.

```{r}
con_dvdrental <- dbConnect(odbc::odbc(), "PostgreSQL AWS dvdrental")
dbListTables(con_dvdrental)
dbListFields(con_dvdrental, 'customer')
```

```{sql, connection=con_dvdrental}
SELECT first_name, last_name, email
  FROM customer
  WHERE first_name = 'Terry'
    AND last_name  = 'Grissom';
```

```{sql, connection=con_dvdrental}
SELECT *
  FROM customer
  WHERE first_name = 'Terry'
    AND last_name  = 'Grissom';
```

### Problem 2

It's another data at the DVD rental store and your manager needs assistance to see each of the unique category IDs. Write a query that will return results showing all of the unique category IDs.

```{r}
dbListTables(con_dvdrental)
dbListFields(con_dvdrental, 'category')
```

```{sql, connection=con_dvdrental}
SELECT DISTINCT category_id
  FROM category;
```

For my own curiosity...
```{sql, connection=con_dvdrental}
SELECT DISTINCT category_id, name
  FROM category
  ORDER BY category_id ASC;
```

We're done with this database for now.
```{r}
dbDisconnect(con_dvdrental)
```

### Problem 3

```{r}
con_petshop <- dbConnect(odbc::odbc(), "PostgreSQL AWS petshop")
```

There should be nothing in the _petshop_ database now.

```{sql, connection=con_petshop}
SELECT * FROM dogs;
```

#### Problem 3, Part 1

Now you're an employee at the pet shop, and a new dog has been added to your pet inventory. Write a query to add the new dog using the following information:

* Name -- Beethoven
* Age -- 2
* Weight -- 22

```{sql, connection=con_petshop}
INSERT INTO dogs(
  name, age, weight
) VALUES (
  'Beethoven', 2, 22
);
```

```{sql, connection=con_petshop}
SELECT * FROM dogs;
```

Notice the the `dog_id` for Beethoven is 3.

#### Problem 3, Part 2

After adding the new dog, you realize the weight data is wrong, and it should be 32 instead of 22. Write a query to update the weight for Beethoven.

```{sql, connection=con_petshop}
UPDATE dogs
  SET   weight = 32
  WHERE name   = 'Beethoven';
```

```{sql, connection=con_petshop}
SELECT * FROM dogs;
```

It worked.

#### Probelm 3, Part 3

Still at the pet shop, you are informed that Beethoven has been sold. Write a query to remove Beethoven's record from the `dogs` table.

```{sql, connection=con_petshop}
SELECT *
  FROM dogs
  WHERE name = 'Beethoven';
```

```{sql, connection=con_petshop}
DELETE FROM dogs
  WHERE name = 'Beethoven';
```

```{sql, connection=con_petshop}
SELECT * FROM dogs;
```

The record deletion worked.

Disconnection from the _petshop_ database.
```{r}
dbDisconnect(con_petshop)
```


```
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
#  @##==---==##@##==---==##@    EXTRA  :  SPACE    @##==---==##@##==---==##@  #
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
#  @##==---==##@##==---==##@    EXTRA  :  SPACE    @##==---==##@##==---==##@  #
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
```
