---
title: "Chapter 6"
author: "Nicolas May"
date: "1/5/2019"
output: 
  html_document:
    theme: readable
    highlight: zenburn
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load useful libraries.
```{r libraries}
library(DBI)
library(odbc)
library(RPostgres)
```

# Chapter Sections

Connect to the _dvdrental_ database.

```{r}
con_dvdrental <- dbConnect(odbc::odbc(), "PostgreSQL AWS dvdrental",
                           bigint = "numeric")
```

## Group Data with the GROUP BY Clause

An example using the `GROUP BY` clause:
```
SELECT customer_id, SUM(amount)
  FROM payment
  GROUP BY customer_id;
```

```{sql connection=con_dvdrental}
SELECT customer_id, SUM(amount)
  FROM payment
  GROUP BY customer_id
  ORDER BY SUM(amount) DESC;
```

```{r}
customer_sums <-
  dbGetQuery(con_dvdrental,
             "SELECT customer_id, SUM(amount)
                FROM payment
                GROUP BY customer_id
                ORDER BY sum DESC;")
library(ggplot2)
ggplot(customer_sums, aes(x = sum)) +
  geom_histogram(binwidth = 5, na.rm = TRUE) +
  scale_x_continuous(limits = c(0L, 250L))
dbGetQuery(con_dvdrental,
           "SELECT AVG(sub.sum) 
              FROM (
                SELECT customer_id, SUM(amount) as sum
                  FROM payment
                  GROUP BY customer_id
              ) sub;")
```

```{sql connection=con_dvdrental}
SELECT AVG(sub.sum)
  FROM (
    SELECT customer_id, SUM(amount) as sum
      FROM payment
      GROUP BY customer_id
  ) sub;
```

### Exercise

#### Task 1

Write a query that returns the number of films by rating.

```{sql connection=con_dvdrental}
SELECT rating, COUNT(*)
  FROM film
  GROUP BY rating
  ORDER BY rating;
```


## The HAVING Clause

An example using the `HAVING` filter:
```
SELECT customer_id, SUM(amount)
  FROM payment
  GROUP BY customer_id
  HAVING SUM(amount) > 200;
```

`HAVING` vs. `WHERE` -- `WHERE` is used to filter rows _*BEFORE*_ a `GROUP BY` clause, whereas `HAVING` is used to filter rows _*AFTER*_ a `GROUP BY` clause.

An example with both `WHERE` and `HAVING`:
```
SELECT customer_id, SUM(amount)
  FROM payment
  WHERE customer_id != 10
  GROUP BY customer_id
  HAVING SUM(amount) > 200;
```

```{sql connection=con_dvdrental}
SELECT customer_id, SUM(amount)
  FROM payment
  GROUP BY customer_id
  HAVING SUM(amount) > 100
  ORDER BY SUM(AMOUNT) ASC;
```

```{sql connection=con_dvdrental}
SELECT customer_id, SUM(amount)
  FROM payment
  WHERE customer_id BETWEEN 25 AND 50
  GROUP BY customer_id
  HAVING SUM(amount) >= 125.00
  ORDER BY customer_id;
```

### Exercise

#### Task 1

Write a query that returns the number of films by rating with the rating 'G'.

```{sql connection=con_dvdrental}
SELECT rating, COUNT(*)
  FROM film
  GROUP BY rating
  HAVING rating = 'G';
```

Here's a better exercise/task:
```{sql connection=con_dvdrental}
SELECT rating, COUNT(*)
  FROM film
  WHERE length > 150
  GROUP BY rating
  HAVING rating IN ('G', 'PG');
```

## PostgreSQL Aliases and the LIMIT Clause

Aliases with `AS`:
```{sql connection=con_dvdrental}
SELECT first_name AS f_name
  FROM customer
  WHERE first_name LIKE 'Ad%';
```

Limiting results returned with `LIMIT`:
```{sql connection=con_dvdrental}
SELECT *
  FROM customer
  ORDER BY customer_id DESC
  LIMIT 5;
```

Using `OFFSET` with `ORDER BY`:
```{sql connection=con_dvdrental}
SELECT *
  FROM customer
  ORDER BY customer_id ASC
  LIMIT 5
  OFFSET 3;
```

```{sql connection=con_dvdrental}
SELECT customer_id AS id, first_name AS fn, last_name as ln
  FROM customer
  ORDER BY customer_id
  LIMIT 7 OFFSET 13;
```

```{r}
dbDisconnect(con_dvdrental)
```


```
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
#  @##==---==##@##==---==##@    EXTRA  :  SPACE    @##==---==##@##==---==##@  #
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
#  @##==---==##@##==---==##@    EXTRA  :  SPACE    @##==---==##@##==---==##@  #
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
```



