---
title: "Chapter4"
author: "Nicolas May"
date: "January 3, 2019"
output: 
  html_document:
    theme: readable
    highlight: zenburn
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load useful libraries.
```{r libraries}
library(DBI)
library(odbc)
library(RPostgres)
```


# Chapter Sections

Connect to the _dvdrental_ database.
```{r}
con_dvdrental <- dbConnect(odbc::odbc(), "PostgreSQL AWS dvdrental")
```


## Query Rows of Data Using the BETWEEN Operator

The template for using the `BETWEEN` operator is:
```
SELECT column_name_x
  FROM table_name
  WHERE column_name_y
  BETWEEN low_value
    AND high_value;
```

Here's a working example:
```{sql connection=con_dvdrental}
SELECT customer_id, first_name, last_name, email
  FROM customer
  WHERE customer_id
    BETWEEN 1 AND 10;
```

### Exercise

#### Task 1

Write a query that selects the `customer_id` column fromt the `payment` table where the value range of the amount column is between 10.00 and 12.00.

```{r}
dbListFields(con_dvdrental, "payment")
```

```{sql connection=con_dvdrental}
SELECT customer_id, amount
  FROM payment
  WHERE amount
    BETWEEN 10.00 AND 12.00
  ORDER BY amount DESC; -- added an ORDER BY
```


## Introduction to NOT BETWEEN

Working example:
```{sql connection=con_dvdrental}
SELECT customer_id, first_name, last_name
  FROM customer
  WHERE customer_id
    NOT BETWEEN 1 AND 10
  ORDER BY customer_id;
```

### Exercise

#### Task 1

Write a query that selects the `replacement_cost` and `title` columns from the `film` table where the values are outside the range of 1.00 through 15.00.

```{sql connection=con_dvdrental}
SELECT replacement_cost, title
  FROM film
  WHERE replacement_cost
    NOT BETWEEN 1.00 AND 15.00
  ORDER BY replacement_cost DESC;
```


## Search for a Specified Pattern in a Column Using the LIKE Operator

The `LIKE` operator is useful for string pattern matching. Examples:

* Values in the `first_name` column that start with 'A'
* Values in the `last_name` column that end with 'e'

The `LIKE` operator uses wildcards:
1. `%` - zero, one, or multiple characters (like regex `*`)
2. `_` - one character (like regex `?`)

Here's a template for using the `LIKE` operator:
```
SELECT first_name, last_name
  FROM customer
  WHERE first_name
    LIKE 'A%';      -- matches any last name starting with 'A'
```

Let's try a few examples.

```{sql connection=con_dvdrental}
SELECT first_name, last_name
  FROM customer
  WHERE first_name
    LIKE 'A%';
```

```{sql connection=con_dvdrental}
SELECT first_name, last_name
  FROM customer
  WHERE last_name
    LIKE '%e';
```

```{sql connection=con_dvdrental}
SELECT first_name, last_name
  FROM customer
  WHERE first_name
    LIKE '_d%';
```

### Exercise

#### Task 1

Write a query that selects all columns from the actor table but returns rows where values in the `last_name` column end with 'da'.

```{r}
dbListFields(con_dvdrental, "actor")
```

```{sql connection=con_dvdrental}
SELECT *
  FROM actor
  WHERE last_name LIKE '%da';
```

From the PG documentation, I see we can also use POSIX regex with the `~` operator.

```{sql connection=con_dvdrental}
SELECT *
  FROM actor
  WHERE last_name ~ '^C.*e$'; -- POSIX regex, starts w/ 'C', ends with 'e'
```


## Match a Value to a List of Values with the IN Operator

An example of use of the `IN` operator:
```
SELECT first_name, last_name
  FROM customer
  WHERE first_name
    IN ('John', 'Mary', 'Timothy');
```

```{sql connection=con_dvdrental}
SELECT first_name, last_name
  FROM customer
  WHERE first_name
    IN ('Mary', 'James');
```

### Exercise 

#### Task 1

Write a query that selects the `first_name` and `last_name` columns from the actor table but returns rows where values in the `first_name` column are equal to Morgan, Julia, or Kenneth.

```{sql connection=con_dvdrental}
SELECT first_name, last_name
  FROM actor
  WHERE first_name
    IN ('Morgan', 'Julia', 'Kenneth');
```


## Assignment

### Problem 1

It's another day at the DVD rental store, and your manager needs you to query the database for all the customer by `customer_id` 1 through 10.

```{sql connection=con_dvdrental}
SELECT *
  FROM customer
  WHERE customer_id
    BETWEEN 1 AND 10;
```

### Problem 2

Later in the day, your manager asks you to query all the customer from the database _except_ those with `customer_id` 1 through 10.

```{sql connection=con_dvdrental}
SELECT *
  FROM customer
  WHERE customer_id
    NOT BETWEEN 1 AND 10
  ORDER BY customer_id ASC;
```

### Problem 3

It's the next day at the DVD rental store, and a customer is having trouble finding a particular film. They tell you they know the film title starts with the characters 'Tru', but they're unsure how to spell the rest of the title. Write a query that will return the films matching this description.

```{r}
dbListFields(con_dvdrental, "film")
```

```{sql connection=con_dvdrental}
SELECT title, release_year, rental_rate, length, rating
  FROM film
  WHERE title LIKE 'Tru%';
```

### Problem 4

Right after helping your last customer, your coworker asks you for help to audit all the films that have a certain rating. Write a query for the film table that will return rows where the rating is G, PG-13, or R.

```{sql connection=con_dvdrental}
SELECT title, length, rating
  FROM film
  WHERE rating 
    IN ('G', 'PG-13', 'R')
  ORDER BY length DESC;
```

### Disconnect from database

Disconnect from _dvdrental_ database.
```{r}
dbDisconnect(con_dvdrental)
```


```
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
#  @##==---==##@##==---==##@    EXTRA  :  SPACE    @##==---==##@##==---==##@  #
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
#  @##==---==##@##==---==##@    EXTRA  :  SPACE    @##==---==##@##==---==##@  #
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
```
