---
title: "Chapter4"
author: "Nicolas May"
date: "January 3, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load useful libraries.
```{r libraries}
library(DBI)
library(odbc)
library(RPostgres)
```


# Chapter Sections

Connect to the _dvdrental_ database.
```{r}
con_dvdrental <- dbConnect(odbc::odbc(), "PostgreSQL AWS dvdrental")
```


## Query Rows of Data Using the BETWEEN Operator

The template for using the `BETWEEN` operator is:
```
SELECT column_name_x
  FROM table_name
  WHERE column_name_y
  BETWEEN low_value
    AND high_value;
```

Here's a working example:
```{sql connection=con_dvdrental}
SELECT customer_id, first_name, last_name, email
  FROM customer
  WHERE customer_id
  BETWEEN 1
    AND 10;
```

### Exercise

#### Task 1

Write a query that selects the `customer_id` column fromt the `payment` table where the value range of the amount column is between 10.00 and 12.00.

```{r}
dbListFields(con_dvdrental, "payment")
```

```{sql connection=con_dvdrental}
SELECT customer_id, amount
  FROM payment
  WHERE amount
  BETWEEN 10.00
    AND 12.00
  ORDER BY amount DESC; -- added an ORDER BY
```


## Introduction to NOT BETWEEN

Working example:
```{sql connection=con_dvdrental}
SELECT customer_id, first_name, last_name
  FROM customer
  WHERE customer_id
  NOT BETWEEN 1
    AND 10
  ORDER BY customer_id;
```

### Exercise

#### Task 1

Write a query that selects the `replacement_cost` and `title` columns from the `film` table where the values are outside the range of 1.00 through 15.00.

```{sql connection=con_dvdrental}
SELECT replacement_cost, title
  FROM film
  WHERE replacement_cost
  NOT BETWEEN 1.00
    AND 15.00
  ORDER BY replacement_cost DESC;
```


## Search for a Specified Pattern in a Column Using the LIKE Operator



Disconnect from _dvdrental_ database.
```{r}
dbDisconnect(con_dvdrental)
```




```
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
#  @##==---==##@##==---==##@    EXTRA  :  SPACE    @##==---==##@##==---==##@  #
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
#  @##==---==##@##==---==##@    EXTRA  :  SPACE    @##==---==##@##==---==##@  #
#@##==---==##@   @##==---==##@    #==-- --==#    @##==---==##@   @##==---==##@#
##==---==##@   #   @##==---==##@    #==-==#    @##==---==##@   #   @##==---==##
#=---==##@    #=#    @##==---==##@    #=#    @##==---==##@    #=#    @##==---=#
#--==##@    #==-==#    @##==---==##@   #   @##==---==##@    #==-==#    @##==--#
#==##@    #==-- --==#    @##==---==##@   @##==---==##@    #==-- --==#    @##==#
###@    #==--  :  --==#    @##==---==##@##==---==##@    #==--  :  --==#    @###
```
